Dominance Regions for Rank Two Cluster Algebras

### 2022-04-07 dominance_regions.tex
Superscripts section 4

  Important examples of pointed elements are the \emph{cluster monomials} of $\cA(b,c)$, i.e. the elements that can be written in the form $x_k^{a_k}x_{k+1}^{a_{k+1}}$ for some $k\in\ZZ$ and $a_k,a_{k+1}$ non-negative integers.
  Indeed, by [someresult of Qin] cluster monomials are part of any pointed basis in any cluster algebra.
  \say{Check if this needs full rank}

@@

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We will therefore assume that $bc\ge4$.
  In this case we give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b a_0 ,c a_1)$, $a_0,a_1\in\ZZ$, inside its \emph{dominance polygon} $\cP(\lambda)$.
  %In this case, we compute the possible transitions between any two pointed bases.
  %The collection of $\bfg$-vectors which can appear in the deformation of a pointed basis is the \emph{dominance region}.
  \begin{theorem}
    \label{th:dominance inequalities}
    The dominance polygon $\cP(\lambda)$ pointed at $\lambda=(\lambda_0,\lambda_1)$ is the region consisting of those $\mu\in\RR^2$ satisfying $\mu_0 \leq \lambda_0, \mu_1 \geq\lambda_1$, and the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.8}
      \[
        \begin{array}{rcccl}
          0 & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \leq & \max\left(-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1,0\right)
          \\
          \min\left(-c\lambda_0,0\right) & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) & \leq & 0
          \\
          0 & \leq &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \leq & \max\left(\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1,0\right)
          \\
          \min\left(b \lambda_1,0\right) & \leq & (\mu_0-\lambda_0) - \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \leq & 0
        \end{array}
      \]
    }
  \end{theorem}

  A number of bases for rank two cluster algebras with various defining properties have been constructed.
  A natural starting basis is the so-called \emph{greedy basis} in which all $c^\lambda_{a_1,a_2}$ are non-negative and these coefficients are chosen to be minimal.
  It is known that this choice coincides with the theta basis \cite{GHKK, CGMMRSW}.
  The \emph{standard monomial bases} of $\cA(b,c)$ come from natural identifications with the subalgebras of $\ZZ[x_0^{\pm1},x_1^{\pm1}]$ generated by $x_{k-1}$, $x_k$, $x_{k+1}$, and $x_{k+2}$ for $k\in\ZZ$.
  However, these bases depend on $k$ and the triangular basis is defined by a triangular transition with the stipulation that the basis does not depend on the choice of starting cluster.
  It would be useful to explicitly (or combinatorially) compute the transitions between these bases.
  However, explicit calculations reveal that the $\bfg$-vectors actually appearing in these transitions are far smaller than the entire dominance polygon.

@@

  We record here a few useful calculations relating to the tropical transformations $\phi_k$.
  First observe the following explicit expression for $\phi_2$:
  \begin{equation}
    \label{eq:forward two step mutation}
    \phi_2(\lambda)
    =
    \begin{cases}
      \big((bc-1)\lambda_0+b\lambda_1, -c\lambda_0-\lambda_1\big) & \text{if $\lambda_0\ge 0$ and $c\lambda_0+\lambda_1\ge 0$;}\\
      (-\lambda_0, -c\lambda_0-\lambda_1) & \text{if $\lambda_0\ge 0$ and $c\lambda_0+\lambda_1<0$;}\\
      (-\lambda_0+b\lambda_1, -\lambda_1) & \text{if $\lambda_0<0$ and $\lambda_1\ge 0$;}\\
      (-\lambda_0,-\lambda_1) & \text{if $\lambda_0<0$ and $\lambda_1<0$;}
    \end{cases}
  \end{equation}
  %\begin{equation}
  %  \label{eq:backward two step mutation}
  %  \phi_{-2}(\lambda)
  %  =
  %  \begin{cases}
  %    \big(-\lambda_0-b\lambda_1, c\lambda_0+(bc-1)\lambda_1\big) & \text{if $\lambda_1\le 0$ and $\lambda_0+b\lambda_1\le 0$;}\\
  %    (-\lambda_0-b\lambda_1, -\lambda_1) & \text{if $\lambda_1\le 0$ and $\lambda_0+b\lambda_1>0$;}\\
  %    (-\lambda_0, c\lambda_0-\lambda_1) & \text{if $\lambda_1>0$ and $\lambda_0\le 0$;}\\
  %    (-\lambda_0,-\lambda_1) & \text{if $\lambda_1>0$ and $\lambda_0>0$.}
  %  \end{cases}
  %\end{equation}

@@

  It will be useful to have explicit expressions for $\phi_k(\lambda)$ for $\lambda\in\cI$.
  \begin{lemma}
    \label{le:imaginary transformations}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have
    \begin{align*}
      \phi_{2j}(\lambda)&=(u_{2j+1,+}\lambda_0+u_{2j,+}\lambda_1,-u_{2j,-}\lambda_0-u_{2j-1,-}\lambda_1);\\
      \phi_{2j+1}(\lambda)&=(-u_{2j+1,+}\lambda_0-u_{2j,+}\lambda_1,u_{2j+2,-}\lambda_0+u_{2j+1,-}\lambda_1).
    \end{align*}
  \end{lemma}
  \begin{proof}
    We work by induction on $j$, the case $j=0$ being clear from the definitions.
    For $\lambda\in\cI$, the action of $\phi_2$ from \eqref{eq:forward two step mutation} can be rewritten as
    \[\phi_2(\lambda)=(u_{3,-}\lambda_0+u_{2,+}\lambda_1,-u_{2,-}\lambda_0-u_{1,+}\lambda_1).\]
    Therefore
    \begin{align*}
      \phi_{2j+2}(\lambda)&=\phi_2\phi_{2j}(\lambda)\\
      &=\big( (u_{3,-}u_{2j+1,+}-u_{2,+}u_{2j,-})\lambda_0+(u_{3,-}u_{2j,+}-u_{2,+}u_{2j-1,-})\lambda_1,\\
      &\qquad -(u_{2,-}u_{2j+1,+}-u_{1,+}u_{2j,-})\lambda_0-(u_{2,-}u_{2j,+}-u_{1,+}u_{2j-1,-})\lambda_1 \big)\\
      &=(u_{2j+3,+}\lambda_0+u_{2j+2,+}\lambda_1, -u_{2j+2,-}\lambda_0-u_{2j+1,-}\lambda_1),
    \end{align*}
    as desired, where the last equality uses \eqref{eq:long Chebyshev recursion} with $i=2j+1$ and $\ell=2$.

    Similarly, using $\phi_1(\lambda)=(-\lambda_0,u_{2,-}\lambda_0+u_{1,+}\lambda_1)$ for $\lambda\in\cI$ together with the basic Chebyshev recursion and the equality $\phi_{2j+1}=\phi_1\phi_2^j$ gives the claimed formula for $\phi_{2j+1}$ from that of $\phi_{2j}$.
  \end{proof}

@@

  Here we explicitly compute the dominance regions $\cP(\lambda)$.
  When $\lambda$ is not imaginary, the dominance region is particularly simple.
  \begin{lemma}
    If $\lambda\in\ZZ^2\setminus\cI$, then $\cP(\lambda)=\{\lambda\}$.
  \end{lemma}
  \begin{proof}
    Any such $\lambda$ is the $\bfg$-vector of a cluster monomial, say $x_k^{a_k}x_{k+1}^{a_{k+1}}$.
    In this case the intersection
    \[\phi_{k-1}^{-1}\cC_{k-1}(\phi_{k-1}\lambda) \cap \phi_{k+1}^{-1}\cC_{k+1}(\phi_{k+1}\lambda)\]
    is precisely $\{\lambda\}$. 
    Indeed, when $k$ is odd the cone $\cC_{k-1}(\phi_{k-1}\lambda)$ lies entirely within a domain of linearity for $\phi_{k-1}^{-1}$.
    In particular, $\phi_{k-1}^{-1}\cC_{k-1}(\phi_{k-1}\lambda)$ is a cone containing $\lambda$ directed away from the origin (with walls parallel to the boundary of the domain of linearity containing $\lambda$).
    On the contrary, when $k$ is odd the cone $\cC_{k+1}(\phi_{k+1}\lambda)$ intersects the domain of linearity for $\phi_{k+1}^{-1}$ containing $\phi_{k+1}\lambda$ (resp. $\phi_{k-1}\lambda$) in a (possibly degenerate) convex quadrilateral with corners at $(0,0)$ and $\phi_{k+1}\lambda$ (resp. $\phi_{k-1}\lambda$).
    In particular, $\phi_{k+1}^{-1}\cC_{k+1}(\phi_{k+1}\lambda)$ (resp. $\phi_{k-1}^{-1}\cC_{k-1}(\phi_{k-1}\lambda)$) is a (possibly degenerate) convex quadrilateral with corners at $(0,0)$ and $\lambda$.
    The result follows.
  \end{proof}
  To compare with Theorem~\ref{th:dominance inequalities}, we consider three cases in which $\lambda\in\ZZ^2\setminus\cI$:
  \begin{itemize}
    \item if $\lambda_0 \le 0$ and $\lambda_1 \ge 0$, the second and fourth inequalities become equations for lines which intersect at $\lambda$;
    \item if $\lambda_0 > 0$ and $\lambda_1 \ge \frac{-bc+\sqrt{bc(bc-4)}}{2b}\lambda_0$, the first and second inequalities become equations for lines which intersect at $\lambda$;
    \item if $\lambda_1 < 0$ and $\lambda_0 \le \frac{bc-\sqrt{bc(bc-4)}}{2c}\lambda_1$, the third and fourth inequalities become equations for lines which intersect at $\lambda$.
  \end{itemize}
  This proves Theorem~\ref{th:dominance inequalities} when $\lambda\notin\cI$.

@@

  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A$ be an invertible $2\times 2$ matrix.
    Under the left action of $A$, say $A\bfx=\bfy$ for $\bfx,\bfy\in\RR^2$, the region inside $\RR^2$ defined by the inequality $\bfr\bullet\bfx \le t$ with $\bfr\in\RR^2$ and $t\in\RR$ is transformed into the region defined by the inequality \[(A^{-T}\bfr)\bullet\bfy\le t.\]
  \end{lemma}
  \begin{proof}
    This is immediate from the equalities
    \[\bfr\bullet\bfx=\bfr^T\bfx=\bfr^TA^{-1}A\bfx=(A^{-T}\bfr)^TA\bfx=(A^{-T}\bfr)\bullet\bfy.\]
  \end{proof}

  \begin{lemma}
    For $k\ge1$ and $\lambda\in\cI$, the region $\phi_k^{-1}\cC_k(\phi_k\lambda)\subset\RR^2$ is the polytope determined by the following inequalities:
    \begin{align}
      \label{ineq:1} u_{k,-}x+u_{k-1,+}y &\le u_{k,-}\lambda_0+u_{k-1,+}\lambda_1;\\
      \label{ineq:2} u_{k+1,-}x+u_{k,+}y &\le u_{k+1,-}\lambda_0+u_{k,+}\lambda_1;\\
      \label{ineq:3} -u_{k-1,-}x+u_{k,+}y &\le u_{k+1,-}\lambda_0+u_{k,+}\lambda_1;\\
      \label{ineq:4} -u_{k-1,-}x-u_{k-2,+}y &\le u_{k+1,-}\lambda_0+u_{k,+}\lambda_1.
    \end{align}

@@

    For $k=2j$, $j>0$, and $\lambda\in\cI$, we observe that $\cC_k(\phi_k\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger)\ x\le u_{2j+1,+}\lambda_0+u_{2j,+}\lambda_1 \qquad\text{and}\qquad (\ddagger)\ -y\le u_{2j,-}\lambda_0+u_{2j-1,-}\lambda_1. \]
    We inductively compute the region $\phi_{2j}^{-1}\cC_{2j}(\phi_{2j}\lambda)$, using Lemma~\ref{le:transformed inequalities} and the equality $\phi_{2j}^{-1}=\phi_2^{-j}$.

    \subsubsection*{Claim:} For $i>0$, $\phi_{2i}^{-1}\cC_{2j}(\phi_{2j}\lambda)$ is the region determined by the inequalities 
    \begin{align*}
      \tag{a} u_{2i,-}x+u_{2i-1,+}y &\le u_{2j,-}\lambda_0+u_{2j-1,+}\lambda_1;\\
      \tag{b} u_{2i+1,-}x+u_{2i,+}y &\le u_{2j+1,-}\lambda_0+u_{2j,+}\lambda_1;\\
      \tag{c} -u_{2i-1,-}x+u_{2i,+}y &\le u_{2j+1,-}\lambda_0+u_{2j,+}\lambda_1;\\
      \tag{d} -u_{2i-1,-}x-u_{2i-2,+}y &\le u_{2j+1,-}\lambda_0+u_{2j,+}\lambda_1.
    \end{align*}

    We see from \eqref{eq:backward two step mutation}, that $\lambda\in\cI$ implies the boundary ray for $\cC_{2j}(\phi_{2j}\lambda)$ corresponding to ($\ddagger$) lies entirely in the region in which $\phi_2^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & -b\\ c & bc-1 \end{array}\right]$ of determinant 1.
    By Lemma~\ref{le:transformed inequalities}, the inequality ($\ddagger$) transforms into the inequality $cx+y\le u_{2j,-}\lambda_0+u_{2j-1,+}\lambda_1$ which corresponds to (a) with $i=1$.
    Similarly, the boundary ray for $\cC_{2j}(\phi_{2j}\lambda)$ corresponding to ($\dagger$) intersects the three domains of linearity in which $\phi_2^{-1}$ acts according to the matrices $\left[ \begin{array}{cc} -1 & -b\\ c & bc-1 \end{array}\right]$, $\left[ \begin{array}{cc} -1 & -b\\ 0 & -1 \end{array}\right]$, $\left[ \begin{array}{cc} -1 & 0\\ 0 & -1 \end{array}\right]$, each of determinant 1.
    By Lemma~\ref{le:transformed inequalities}, the inequality ($\dagger$) can be seen to transform by these into each of inequalities (b), (c), (d) with $i=1$.
    This establishes the base of an induction on $i$.

>>>>

  Important examples of pointed elements are the \emph{cluster monomials} of $\cA(b,c)$, i.e. the elements that can be written in the form $x_k^{\alpha_k}x_{k+1}^{\alpha_{k+1}}$ for some $k\in\ZZ$ and $a_k,a_{k+1}$ non-negative integers.
  Indeed, by [someresult of Qin] cluster monomials are part of any pointed basis in any cluster algebra.
  \say{Check if this needs full rank}

@@

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We will therefore assume that $bc\ge4$.
  In this case we give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside its \emph{dominance polygon} $\cP(\lambda)$.
  %In this case, we compute the possible transitions between any two pointed bases.
  %The collection of $\bfg$-vectors which can appear in the deformation of a pointed basis is the \emph{dominance region}.
  \begin{theorem}
    \label{th:dominance inequalities}
    The dominance polygon $\cP(\lambda)$ pointed at $\lambda=(\lambda_0,\lambda_1)$ is the region consisting of those $\mu\in\RR^2$ satisfying $\mu_0 \leq \lambda_0, \mu_1 \geq\lambda_1$, and the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.8}
      \[
        \begin{array}{rcccl}
          0 & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \leq & \max\left(-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1,0\right)
          \\
          \min\left(-c\lambda_0,0\right) & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) & \leq & 0
          \\
          0 & \leq &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \leq & \max\left(\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1,0\right)
          \\
          \min\left(b \lambda_1,0\right) & \leq & (\mu_0-\lambda_0) - \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \leq & 0
        \end{array}
      \]
    }
  \end{theorem}

  A number of bases for rank two cluster algebras with various defining properties have been constructed.
  A natural starting basis is the so-called \emph{greedy basis} in which all $c_{\alpha_0,\alpha_1}$ are non-negative and these coefficients are chosen to be minimal.
  It is known that this choice coincides with the theta basis \cite{GHKK, CGMMRSW}.
  The \emph{standard monomial bases} of $\cA(b,c)$ come from natural identifications with the subalgebras of $\ZZ[x_0^{\pm1},x_1^{\pm1}]$ generated by $x_{k-1}$, $x_k$, $x_{k+1}$, and $x_{k+2}$ for $k\in\ZZ$.
  However, these bases depend on $k$ and the triangular basis is defined by a triangular transition with the stipulation that the basis does not depend on the choice of starting cluster.
  It would be useful to explicitly (or combinatorially) compute the transitions between these bases.
  However, explicit calculations reveal that the $\bfg$-vectors actually appearing in these transitions are far smaller than the entire dominance polygon.

@@

  We record here a few useful calculations relating to the tropical transformations $\phi_k$.
  First observe the following explicit expression for $\phi_2$:
  \begin{equation}
    \label{eq:forward two step mutation}
    \phi_2(\lambda)
    =
    \begin{cases}
      \big((bc-1)\lambda_0+b\lambda_1, -c\lambda_0-\lambda_1\big) & \text{if $\lambda_0\ge 0$ and $c\lambda_0+\lambda_1\ge 0$;}\\
      (-\lambda_0, -c\lambda_0-\lambda_1) & \text{if $\lambda_0\ge 0$ and $c\lambda_0+\lambda_1<0$;}\\
      (-\lambda_0+b\lambda_1, -\lambda_1) & \text{if $\lambda_0<0$ and $\lambda_1\ge 0$;}\\
      (-\lambda_0,-\lambda_1) & \text{if $\lambda_0<0$ and $\lambda_1<0$;}
    \end{cases}
  \end{equation}
  \begin{equation}
    \label{eq:backward two step mutation}
    \phi_{-2}(\lambda)
    =
    \begin{cases}
      \big(-\lambda_0-b\lambda_1, c\lambda_0+(bc-1)\lambda_1\big) & \text{if $\lambda_1\le 0$ and $\lambda_0+b\lambda_1\le 0$;}\\
      (-\lambda_0-b\lambda_1, -\lambda_1) & \text{if $\lambda_1\le 0$ and $\lambda_0+b\lambda_1>0$;}\\
      (-\lambda_0, c\lambda_0-\lambda_1) & \text{if $\lambda_1>0$ and $\lambda_0\le 0$;}\\
      (-\lambda_0,-\lambda_1) & \text{if $\lambda_1>0$ and $\lambda_0>0$.}
    \end{cases}
  \end{equation}

@@

  It will be useful to have explicit expressions for $\phi_k(\lambda)$ for $\lambda\in\cI$.
  \begin{lemma}
    \label{le:imaginary transformations}
    For $j\in\ZZ$ and $\lambda\in\cI$, we have
    \begin{align*}
      \phi_{2j}(\lambda)&=(u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1,-u_{2j}^-\lambda_0-u_{2j-1}^-\lambda_1);\\
      \phi_{2j+1}(\lambda)&=(-u_{2j+1}^+\lambda_0-u_{2j}^+\lambda_1,u_{2j+2}^-\lambda_0+u_{2j+1}^-\lambda_1).
    \end{align*}
  \end{lemma}
  \begin{proof}
    We work by induction on $j$, the case $j=0$ being clear from the definitions.
    For $\lambda\in\cI$, the action of $\phi_2$ from \eqref{eq:forward two step mutation} can be rewritten as
    \[\phi_2(\lambda)=(u_3^-\lambda_0+u_2^+\lambda_1,-u_2^-\lambda_0-u_1^+\lambda_1).\]
    Therefore
    \begin{align*}
      \phi_{2j+2}(\lambda)&=\phi_2\phi_{2j}(\lambda)\\
      &=\big( (u_3^- u_{2j+1}^+-u_2^+u_{2j}^-)\lambda_0+(u_3^-u_{2j}^+-u_2^+u_{2j-1}^-)\lambda_1,-(u_2^-u_{2j+1}^+-u_1^+u_{2j}^-)\lambda_0-(u_2^-u_{2j}^+-u_1^+u_{2j-1}^-)\lambda_1 \big)\\
      &=(u_{2j+3}^+\lambda_0+u_{2j+2}^+\lambda_1, -u_{2j+2}^-\lambda_0-u_{2j+1}^-\lambda_1),
    \end{align*}
    as desired, where the last equality uses \eqref{eq:long Chebyshev recursion} with $i=2j+1$ and $\ell=2$.

    Similarly, using $\phi_1(\lambda)=(-\lambda_0,u_2^-\lambda_0+u_1^+\lambda_1)$ for $\lambda\in\cI$ together with the basic Chebyshev recursion and the equality $\phi_{2j+1}=\phi_1\phi_2^j$ gives the claimed formula for $\phi_{2j+1}$ from that of $\phi_{2j}$.
  \end{proof}

@@

  Here we explicitly compute the dominance regions $\cP(\lambda)$.
  When $\lambda$ is not imaginary, the dominance region is particularly simple.
  \begin{lemma}
    If $\lambda\in\ZZ^2\setminus\cI$, then $\cP(\lambda)=\{\lambda\}$.
  \end{lemma}
  \begin{proof}
    Any such $\lambda$ is the $\bfg$-vector of a cluster monomial, say $x_k^{\alpha_k}x_{k+1}^{\alpha_{k+1}}$.
    In this case the intersection
    \[\phi_{k-1}^{-1}\cC_{k-1}(\phi_{k-1}\lambda) \cap \phi_{k+1}^{-1}\cC_{k+1}(\phi_{k+1}\lambda)\]
    is precisely $\{\lambda\}$. 
    Indeed, when $k$ is odd the cone $\cC_{k-1}(\phi_{k-1}\lambda)$ lies entirely within a domain of linearity for $\phi_{k-1}^{-1}$.
    In particular, $\phi_{k-1}^{-1}\cC_{k-1}(\phi_{k-1}\lambda)$ is a cone containing $\lambda$ directed away from the origin (with walls parallel to the boundary of the domain of linearity containing $\lambda$).
    On the contrary, when $k$ is odd the cone $\cC_{k+1}(\phi_{k+1}\lambda)$ intersects the domain of linearity for $\phi_{k+1}^{-1}$ containing $\phi_{k+1}\lambda$ (resp. $\phi_{k-1}\lambda$) in a (possibly degenerate) convex quadrilateral with corners at $(0,0)$ and $\phi_{k+1}\lambda$ (resp. $\phi_{k-1}\lambda$).
    In particular, $\phi_{k+1}^{-1}\cC_{k+1}(\phi_{k+1}\lambda)$ (resp. $\phi_{k-1}^{-1}\cC_{k-1}(\phi_{k-1}\lambda)$) is a (possibly degenerate) convex quadrilateral with corners at $(0,0)$ and $\lambda$.
    The result follows.
  \end{proof}
  To compare with Theorem~\ref{th:dominance inequalities}, we consider three cases in which $\lambda\in\ZZ^2\setminus\cI$:
  \begin{itemize}
    \item if $\lambda_0 \le 0$ and $\lambda_1 \ge 0$, the second and fourth inequalities become equations for lines which intersect at $\lambda$;
    \item if $\lambda_0 > 0$ and $\lambda_1 \ge \frac{-bc+\sqrt{bc(bc-4)}}{2b}\lambda_0$, the first and second inequalities become equations for lines which intersect at $\lambda$;
    \item if $\lambda_1 < 0$ and $\lambda_0 \le \frac{bc-\sqrt{bc(bc-4)}}{2c}\lambda_1$, the third and fourth inequalities become equations for lines which intersect at $\lambda$.
  \end{itemize}
  This proves Theorem~\ref{th:dominance inequalities} when $\lambda\notin\cI$.

@@

  \begin{lemma}
    \label{le:transformed inequalities}
    Let $A$ be an invertible $2\times 2$ matrix.
    Under the left action of $A$, say $A\mu=\mu'$ for $\mu,\mu'\in\RR^2$, the region inside $\RR^2$ defined by the inequality $\bfr\bullet\mu \le t$ with $\bfr\in\RR^2$ and $t\in\RR$ is transformed into the region defined by the inequality \[(A^{-T}\bfr)\bullet\mu'\le t.\]
  \end{lemma}
  \begin{proof}
    This is immediate from the equalities
    \[\bfr\bullet\mu=\bfr^T\mu=\bfr^TA^{-1}A\mu=(A^{-T}\bfr)^TA\mu=(A^{-T}\bfr)\bullet\mu'.\]
  \end{proof}

  \begin{lemma}
    For $k\ge1$ and $\lambda\in\cI$, the region $\phi_k^{-1}\cC_k(\phi_k\lambda)\subset\RR^2$ is the polytope determined by the following inequalities:
    \begin{align}
      \label{ineq:1} u_k^-\mu_0+u_{k-1}^+\mu_1 &\le u_k^-\lambda_0+u_{k-1}^+\lambda_1;\\
      \label{ineq:2} u_{k+1}^-\mu_0+u_k^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1;\\
      \label{ineq:3} -u_{k-1}^-\mu_0+u_k^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1;\\
      \label{ineq:4} -u_{k-1}^-\mu_0-u_{k-2}^+\mu_1 &\le u_{k+1}^-\lambda_0+u_k^+\lambda_1.
    \end{align}

@@

    For $k=2j$, $j>0$, and $\lambda\in\cI$, we observe that $\cC_k(\phi_k\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger)\ x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1 \qquad\text{and}\qquad (\ddagger)\ -y\le u_{2j}^-\lambda_0+u_{2j-1}^-\lambda_1. \]
    We inductively compute the region $\phi_{2j}^{-1}\cC_{2j}(\phi_{2j}\lambda)$, using Lemma~\ref{le:transformed inequalities} and the equality $\phi_{2j}^{-1}=\phi_2^{-j}$.

    \subsubsection*{Claim:} For $i>0$, $\phi_{2i}^{-1}\cC_{2j}(\phi_{2j}\lambda)$ is the region determined by the inequalities 
    \begin{align*}
      \tag{a} u_{2i}^-x+u_{2i-1}^+y &\le u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1;\\
      \tag{b} u_{2i+1}^-x+u_{2i}^+y &\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1;\\
      \tag{c} -u_{2i-1}^-x+u_{2i}^+y &\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1;\\
      \tag{d} -u_{2i-1}^-x-u_{2i-2}^+y &\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1.
    \end{align*}

    We see from \eqref{eq:backward two step mutation}, that $\lambda\in\cI$ implies the boundary ray for $\cC_{2j}(\phi_{2j}\lambda)$ corresponding to ($\ddagger$) lies entirely in the region in which $\phi_2^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & -b\\ c & bc-1 \end{array}\right]$ of determinant 1.
    By Lemma~\ref{le:transformed inequalities}, the inequality ($\ddagger$) transforms into the inequality $cx+y\le u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1$ which corresponds to (a) with $i=1$.
    Similarly, the boundary ray for $\cC_{2j}(\phi_{2j}\lambda)$ corresponding to ($\dagger$) intersects the three domains of linearity in which $\phi_2^{-1}$ acts according to the matrices $\left[ \begin{array}{cc} -1 & -b\\ c & bc-1 \end{array}\right]$, $\left[ \begin{array}{cc} -1 & -b\\ 0 & -1 \end{array}\right]$, $\left[ \begin{array}{cc} -1 & 0\\ 0 & -1 \end{array}\right]$, each of determinant 1.
    By Lemma~\ref{le:transformed inequalities}, the inequality ($\dagger$) can be seen to transform by these into each of inequalities (b), (c), (d) with $i=1$.
    This establishes the base of an induction on $i$.

    Assuming the inequalities (a)-(d) hold for $i$, we apply Lemma~\ref{le:transformed inequalities} with $\phi_2^{-1}$.
    Both of the boundary rays corresponding to the inequalities (a) and (d) lies entirely in the region where $\phi_2^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & -b\\ c & bc-1 \end{array}\right]$, also the boundary segment corresponding to (b) intersects this region.
    Thus applying Lemma~\ref{le:transformed inequalities} to (a) gives the inequality 
    \[u_{2i+2}^-x+u_{2i+1}^+y=(u_3^+u_{2i}^--u_2^-u_{2i-1}^+)x+(u_2^+u_{2i}^--u_1^-u_{2i-1}^+)y\le u_{2j}^-\lambda_0+u_{2j-1}^+\lambda_1,\]
    which is the inequality (a) for $i+1$ by Lemma~\ref{eq:long Chebyshev recursion}; while applying this to (d) gives the inequality 
    \[-u_{2i+1}^-x-u_{2i}^+y=(-u_3^+u_{2i-1}^-+u_2^-u_{2i-2}^+)x+(-u_2^+u_{2i-1}^-+u_1^-u_{2i-2}^+)y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which is the inequality (d) for $i+1$ again by Lemma~\ref{eq:long Chebyshev recursion}; finally applying this to (b) gives the inequality 
    \[u_{2i+3}^-x+u_{2i+2}^+y=(u_3^+u_{2i+1}^--u_2^-u_{2i}^+)x+(u_2^+u_{2i+1}^--u_1^-u_{2i}^+)y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which is the inequality (b) for $i+1$.
    Similarly, the boundary segment corresponding to (c) lies entirely in the region where $\phi_2^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & 0\\ c & -1 \end{array}\right]$.
    Thus applying Lemma~\ref{le:transformed inequalities} to (c) gives the inequality 
    \[-u_{2i+1}^-x-u_{2i}^+y=(u_1^+u_{2i-1}^--u_2^-u_{2i}^+)x+(-u_0^+u_{2i-1}^--u_1^-u_{2i}^+)y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which is the inequality (d) for $i+1$ by Lemma~\ref{eq:long Chebyshev recursion}, in particular we see that the segment determined by (c) and the ray determined by (d) align in the image.
    Lastly, the boundary segment corresponding to (b) also intersects the regions where $\phi_2^{-1}$ acts according to the matrices $\left[ \begin{array}{cc} -1 & -b\\ 0 & -1 \end{array}\right]$ and $\left[ \begin{array}{cc} -1 & 0\\ 0 & -1 \end{array}\right]$ respectively.
    Applying Lemma~\ref{le:transformed inequalities} to (b) with the first matrix gives the inequality 
    \[-u_{2i+1}^-x+u_{2i+2}^+y=(-u_1^+u_{2i+1}^-+u_0^-u_{2i}^+)x+(u_2^+u_{2i+1}^--u_1^-u_{2i,+})y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which is the inequality (c) for $i+1$ by Lemma~\ref{eq:long Chebyshev recursion}, while applying Lemma~\ref{le:transformed inequalities} to (b) with the second matrix gives the inequality 
    \[-u_{2i+1}^-x-u_{2i}^+y\le u_{2j+1}^-\lambda_0+u_{2j}^+\lambda_1,\]
    which again reproduces the inequality (d) and aligns with the previous segment and ray in the image.
    This completes the induction on $i$, proving the Claim and the result for $k$ even.

    For $k=2j+1$, $j>0$, and $\lambda\in\cI$, we get $\cC_{2j+1}(\phi_{2j+1}\lambda)\subset\RR^2$ is given by the inequalities 
    \[ (\dagger')\ -x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1 \qquad\text{and}\qquad (\ddagger')\ y\le u_{2j+2}^-\lambda_0+u_{2j+1}^-\lambda_1.\]
    Using that $\phi_{2j+1}^{-1}=\phi_2^{-j}\phi_1^{-1}$, we compute the image inductively as above.
    From \eqref{eq:forward mutation 1} and Lemma~\ref{le:imaginary stability}, we see that the boundary ray for $\cC_{2j+1}(\phi_{2j+1}\lambda)$ corresponding to ($\dagger'$) lies entirely in the region in which $\phi_1^{-1}$ acts according to the matrix $\left[ \begin{array}{cc} -1 & 0\\ c & 1 \end{array}\right]$ of determinant $-1$.
    Thus applying Lemma~\ref{le:transformed inequalities}, the inequality ($\dagger'$) is transformed by $\phi_1^{-1}$ into the inequality $x\le u_{2j+1}^+\lambda_0+u_{2j}^+\lambda_1$.
    The boundary ray corresponding to ($\ddagger'$) intersects both domains of linearity for $\phi_1^{-1}$ and thus produces the inequalities
    \[ cx+y\le u_{2j+2}^+\lambda_0+u_{2j+1}^+\lambda_1 \qquad\text{and}\qquad y\le u_{2j+2}^+\lambda_0+u_{2j+1}^+\lambda_1.\]
